---
output:
  pdf_document:
    latex_engine: pdflatex
    keep_tex: true
    template: "drive:/path/to/template.tex"
---

```{r all_chunks, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      error = FALSE, 
                      warning = FALSE, 
                      message = FALSE, 
                      include = TRUE, 
                      fig.path = 'drive:/path/to/figs')
```

```{r label = "load-pkgs"}
packages <- c("ggplot2", "latex2exp", "tidyverse", "patchwork")
xfun::pkg_attach(packages, install = TRUE, message = FALSE)
```

\pagestyle{empty}
# Single Variable Linear Regression

The data is from the dataset contained in Table 2 on page 7 of Unit 11. It records timber strength *y1*, specific gravity *x1*, and moisture content *x2*.

In the plots, one data point hurt my eyes (!) so I chopped it (edit out *%>% filter(!(x1 == 0.499))* if you want it back). I know that I should either leave it in or run the analysis both with and without it but I leave that until I've got less time.

```{r label = "data", echo = FALSE}
# import data
data.df <- read.csv('o:/m248/data/csvfiles/timber_strength.csv', 
                    fileEncoding = "UTF-8-BOM") %>%
  rename(
    y1 = "Strength", 
    x1 = "Specific.gravity", 
    x2 = "Moisture.content"
  ) %>% 
  filter(!(x1 == 0.499))
# data.df

# create linear regression model:
data.lm <- lm(y1 ~ x1, data = data.df)

data.conf <- data.lm %>% 
  predict(interval = 'confidence', level = 0.95) %>% 
  as.data.frame() %>% 
  rename(
    conf.lo = lwr, 
    conf.hi = upr
  )

data.pred <- data.lm %>% 
  predict(interval = 'prediction', level = 0.95) %>% 
  as.data.frame() %>% 
  rename(
    pred.lo = lwr, 
    pred.hi = upr
  )

data.df <- cbind(data.df, data.conf[!names(data.conf) %in% names(data.df)])
data.df <- cbind(data.df, data.pred[!names(data.pred) %in% names(data.df)])

alpha <- round(as.numeric(data.lm$coefficients[1]), 2)
beta <- round(as.numeric(data.lm$coefficients[2]), 2)
```

For the effect of just specific gravity on timber strength the simple linear regression model is given by:
\[
  Y_i = \alpha + \beta x_i + W_i = `r alpha` + `r beta` \, x_i + W_i \nonumber
\]

A scatterplot and diagnostic plots to assess assumptions of (a) zero mean and constant variance of residuals and (b) normal distribution of residuals around fitted values.

```{r label = "plots", echo = FALSE, fig.height = 7}
# alternative scatterplot (prefer this)
p.scat <- ggplot(data = data.df, aes(x = x1)) +
  geom_ribbon(
    aes(ymin = conf.lo, ymax = conf.hi), 
    fill = "bisque", linetype = "dashed", color = "chocolate", alpha = 0.5
    ) +
  geom_ribbon(
    aes(ymin = pred.lo, ymax = pred.hi), 
    fill = "beige", linetype = "dashed", color = "chocolate", alpha = 0.4
    ) +
  geom_line(
    aes(y = fit), color = "chocolate"
    ) +
  geom_point(
    aes(y = y1), color = "darkred"
    ) + 
  annotate(
    "text", x = 0.54, y = 10.8,
    label = paste("Timber strength =", round(coef(data.lm)[1], 2), " + ", abs(round(coef(data.lm)[2], 2)), " x  specific gravity")
  ) +
  labs(
    title = "Scatterplot with regression line and 95% confidence and prediction bands", 
    y = "Timber strength",
    x = "Specific gravity" 
  ) + 
  theme_bw()

# standardised residuals vs fitted values
p.stres_fit <- ggplot(fortify(data.lm, data.df), aes(.fitted, .stdresid)) +
  geom_point(color = "darkred") +
  geom_hline(yintercept = 0, color = "chocolate") +
  labs(
    title="Std residuals vs Fitted values", 
    y="Standardised residuals",
    x="Fitted values" 
  ) + 
  theme_bw()

# normal probability plot (QQ plot)
p.qqplot <- ggplot(data.lm) +
  stat_qq(aes(sample = .stdresid), color = "darkred") +
  geom_abline(color = "chocolate") +
  labs(
    title="Normal probability plot (QQ plot)", 
    y="",
    x="Theoretical quantiles" 
  ) + 
  theme_bw()

# void plot to create vertical gap in plat_layout()
p.gap <- ggplot() + theme_void()

p.scat / p.gap / (p.stres_fit | p.qqplot) +
  plot_layout(heights = c(8, 1, 3))
```
