---
output:
  pdf_document:
    latex_engine: pdflatex
    keep_tex: true
    template: "d:/path/to/template.tex"

# fontsize: 11pt
---

```{r all_chunks, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      error = FALSE, 
                      warning = FALSE, 
                      message = FALSE, 
                      include = TRUE
```

```{r label = "load-pkgs"}
packages <- c("tidyverse", "latex2exp")
xfun::pkg_attach(packages, install = TRUE, message = FALSE)
```
\pagestyle{empty}
# Creating a plot of a function in Rmarkdown using Pgfplots

I am experimenting with the \LaTeX\ package Pgfplots because I haven't found a way to persuade 'Ggplot2' to produce maths-style plots, with the axes going through the origin. Sure, I can mimic the axes with 'geom_vline(xintercept = 0)' and 'geom_hline(yintercept = 0)' but then adding ticks and labels requires major faffing about.

\begin{tikzpicture}
  \begin{axis}[ dtaxis, 
                title = An example plot, 
                xlabel = {$x$}, 
                xtick = {-4,...,4}, 
                ylabel = {$f(x)$}, 
                ]
      \addplot[ domain = -3.75:3.75, 
                samples = 200, 
                smooth, red, 
                ]
                {(1/4)*x^3};
      \addplot[ domain = -3.25:3, 
                samples = 200, 
                smooth, blue, 
                ]
                {x^2 + 2*x + 1};
  \end{axis}
\end{tikzpicture}

This is generated by:
\begin{verbatim}
\begin{tikzpicture}
  \begin{axis}[ dtaxis, 
                title = An example plot, 
                xlabel = {$x$}, 
                xtick = {-4,...,4}, 
                ylabel = {$f(x)$}, 
                ]
      \addplot[ domain = -3.75:3.75, 
                samples = 200, 
                smooth, red, 
                ]
                {(1/4)*x^3};
      \addplot[ domain = -3.25:3, 
                samples = 200, 
                smooth, blue, 
                ]
                {x^2 + 2*x + 1};
  \end{axis}
\end{tikzpicture}
\end{verbatim}

and with this custom style in the TEX file:
\begin{verbatim}
\pgfplotsset{dtaxis/.style = {
          width = 12cm, 
          title style = { at = {(0, 1)}, yshift = 5mm, 
                          font = \normalsize, }, 
          axis lines = middle, 
          enlargelimits = true, 
          axis line style = {thick, black}, 
          label style = {font = \normalsize, }, 
          xlabel style = {anchor = west, }, 
          ylabel style = {anchor = south, }, 
          tick style = {thick, black}, 
          tick label style = {font = \footnotesize}, 
          tick align = outside, 
          major tick length = {0.6mm}, }}
\end{verbatim}
