---
output:
  pdf_document:
    latex_engine: pdflatex
    keep_tex: true
    template: "x:/path/to/template.tex"
---

```{r all_chunks, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      error = FALSE, 
                      warning =FALSE, 
                      message = FALSE, 
                      include = TRUE)
```

```{r label = "load-pkgs"}
install.packages <- c("ggplot2", "latex2exp", "pwr")
library(ggplot2)
library(latex2exp)
library(pwr)
```
\pagestyle{empty}

\textbf{\Huge\textsf{Power and Sample Size}}
\vspace{5mm}

We shall cover power and sample size calculations, using R with the formula given in Unit 09 and then using an R function. First, a reminder what the Handbook says:

\vspace{5mm}
\textbf{\Large\textsf{Power calculation}}
\vspace{-3mm}

Suppose that a sample of size \(n\) is obtained from a population
distributed as \(N(\mu, \sigma^2)\), where \(\sigma^2\) is known, and the test statistic
\[
  Z_1 = \frac{\overline{X}-\mu_0}{\sigma/\sqrt{n}} \nonumber
\]
is to be used in a test of \(H_0: \mu=\mu_0\) with significance level \(\alpha\). Let \(d > 0\).
\begin{itemize}
  \item When the alternative hypothesis is \(H_1: \mu > \mu_0\) and the true value
        of \(\mu\) is \(\mu_0 + d\) then the \textbf{power of the one-sided test} is
        \[
          1 - \Phi \left(q_{(1-\alpha)} - \frac{d}{\sigma/\sqrt{n}}\right) \nonumber
        \]
  \item When the alternative hypothesis is \(H_1: \mu \not= \mu_0\) and the true value
        of \(\mu\) is \(\mu_0 \pm d\), where \(d\) is not small, then the \textbf{power of the two-sided test} is
        \[
          1 - \Phi \left(q_{(1-\frac{\alpha}{2})} - \frac{d}{\sigma/\sqrt{n}}\right) \nonumber
        \]
\end{itemize}

\vspace{5mm}
\textbf{\normalsize\textsf{Power calculation by formula}}
\vspace{-3mm}

If we use the values of parameters given in Computer Activity 31 of Book 2, we can use R to calculate power using the formula given in \textbf{M248}:

```{r label = "man-pwr", echo=TRUE}
# From Computer Book B, CA31
num <- 25
mu0 <- 5
mu1 <- 7
diff <- mu1 - mu0
stdev <- 10
alpha <- 0.05

# Calculation by formula:
z1 <- qnorm(1-alpha)-(diff/(stdev/sqrt(num)))
pwr.man <- 1 - pnorm(z1)
```
This gives power as `r pwr.man`, which is the same result as given by Minitab.

\newpage{}

\textbf{\normalsize\textsf{Power calculation by function}}
\vspace{-3mm}

To calculate power using an R function, we use the \texttt{pwr.norm.test} of the \texttt{pwr} package.

\texttt{pwr.norm.test(d = NULL, n = NULL, sig.level = 0.05, power = NULL, alternative = c("two.sided","less","greater"))}

\begin{tabular}
  {p{5mm} p{25mm} p{75mm}}
    & d               & Effect size is d = (mu1-mu0)sd \\
    & n               & Number of observations \\
    & sig.level       & Significance level (Type I error probability) \\
    & power           & Power of test (1 minus Type II error probability) \\
    & alternative     & One of "two.sided" (default), "greater", or "less" \\
\end{tabular}
It is a \textbf{one-sample test}. We pass the parameter (one of \textbf{d}, \textbf{n}, \textbf{power} or \textbf{sig.level}) that we want to calculate as \texttt{NULL} and it will be calculated from the others. The function returns a list of the arguments (including the computed one) and adds \textbf{Method} and \textbf{Notes} elements.

The specific value that we want can be isolated using, e.g., 
\texttt{var.name\$power}.

```{r label = "r-pwr", echo=TRUE}
# The pwr package uses d = (mu1-mu0)/sd
pwr.r <- pwr.norm.test(
                      power = NULL, 
                      n = num, 
                      d = diff/stdev, 
                      sig.level = alpha, 
                      alternative="greater")$power
```

The power returned by R is `r pwr.r`, which is the same as that from both manual calculation and Minitab.

```{r label = "plt-pwr", echo=FALSE, out.width="85%", fig.align = 'center'}
# Possible code
h1 <- "greater"     # "two.sided" (default), "greater", or "less"
points <- 20
x.name <- "Effect size"   # or "Sample size"
x.max <- 8.0  #3.0  (200 for sample size; 8.0 for diff)
x.num <- seq(1, points, by=1)
y.int <- pwr.norm.test(d=diff/stdev, n=num, sig.level=0.05, alternative=h1)$power
# for plotting against sample size:  x.vals <- 10 * x.num

x.vals <- x.num * (x.max / points)

powers <- vector(length=length(x.vals))
for (i in 1:length(x.vals)) {
  powers[i] <- pwr.norm.test(
    n = num,    #d = x.vals[i],  for plotting against sample size
    d = (x.vals[i])/stdev, # diff/stdev,   for plotting against sample size
    sig.level = alpha, 
    alternative = h1
  )$power
}
power.df <- data.frame(x.vals, powers)

ggplot(power.df) + 
  aes(x=x.vals, y=powers) + 
  geom_line(
    colour = "darkred",
    size = 0.7
  ) + 
  geom_point(
    colour = "darkred",
    size = 0.7
  ) + 
  geom_segment(mapping = aes(x = diff, y = 0), 
               xend = diff, yend = y.int, 
               linetype = "dashed", 
               colour = "darkred", 
               size = 0.5, 
               show.legend = FALSE
  ) + 
  geom_segment(mapping = aes(x = 0, y.int), 
               xend = diff, yend = y.int, 
               linetype = "dashed", 
               colour = "darkred", 
               size = 0.5, 
               show.legend = FALSE
  ) + 
  theme_bw() + 
  labs(
    title = "Power function", 
    # subtitle="", 
    # caption="",  
    y = "Power",
    x = x.name 
  ) +
  scale_x_continuous(
    limits = c(0, x.max),     # x.max), 
    expand = expansion(mult = c(0, 0.01)),  # gap between axis and (0, 0)
    breaks = seq(0, x.max, by = 1)
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.01)), # gap between axis and (0, 0)
    limits = c(0, 1), 
    breaks = seq(0, 1, by = 0.2)
  ) +
  theme(
    panel.grid = element_line(colour = "white"), 
    plot.title = element_text(size=14)
  )
```
\newpage{}

\textbf{\Large\textsf{Sample size calculation}}
\vspace{-3mm}

Suppose that a sample of data is to be collected and one of the tests
described in the previous point is to be performed. Suppose also that
\(n\) is to be chosen so that the power of the test, when the true
underlying mean is \(\mu_0 + d\), is equal to a predetermined value \(\gamma\).
The required sample size is

\begin{itemize}
  \item For a \textbf{one-sided test}:
        \[
          n = \frac{\sigma^2}{d^2} \left( q_{(1-\alpha)} - q_{(1-\gamma)} \right)^2 \nonumber
        \]
  \item For a \textbf{two-sided test}:
        \[
          n = \frac{\sigma^2}{d^2} \left( q_{(1-\frac{\alpha}{2})} - q_{(1-\gamma)} \right)^2 \nonumber
        \]
\end{itemize}


\textbf{\normalsize\textsf{Sample size by formula}}
\vspace{-3mm}

If we again use the values of parameters given in Computer Activity 31 of Book 2, we can use R to calculate sample size using the formula given in \textbf{M248}:

```{r label = "man-size", echo=TRUE}
# From Book 2, CA31
# num <- 25
gamma <- 0.9
mu0 <- 5
mu1 <- 7
diff <- mu1 - mu0
stdev <- 10
alpha <- 0.05

# Calculation by formula:
size <- (stdev^2/diff^2)*(qnorm(1-alpha)-qnorm(1-gamma))^2
```
This gives a sample size of `r ceiling(size)`, which is the same result as given by Minitab.

\textbf{\normalsize\textsf{Sample size by function}}
\vspace{-3mm}

This time we use the same \texttt{pwr} function that we used for the power calculation but by giving the required power 
and giving a \texttt{NULL} value for \texttt{n}, it returns the sample size needed.

```{r label = "r-size", echo=TRUE}
# The pwr package uses d = (mu-mu_0)/sd
pwr.r <- pwr.norm.test(
                      n = NULL, 
                      power = gamma, 
                      d = diff/stdev, 
                      sig.level = alpha, 
                      alternative="greater")$n
```
This, too, gives a sample size of `r ceiling(size)`, which is the same result as given by both formula and Minitab.
