---
output:
  pdf_document:
    latex_engine: pdflatex
    keep_tex: true
    template: "o:/dtma/dt-tma.tex"

fontsize: 11pt

linestretch: 1.0
---

```{r all_chunks, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r label = "load-pkgs", echo =FALSE, error = FALSE, warnings=FALSE, messages = FALSE}

packages <- c(
"tidyverse", "ggrepel", "ggridges", "ggtext", "ggthemes", "ggstatsplot", "ggforce", "rmarkdown", "xaringan", "xaringanthemer", "kableExtra", "shiny", "corrr", "ggdark", "ggsci", "grid", "gridExtra", "showtext", "rcartocolor", "patchwork", "cowplot", "rvest", "jsonlite", "colorspace", "data.table", "latex2exp", "Hmisc", "broom", "reticulate"
)

xfun::pkg_attach(packages, install = TRUE, message = FALSE)

```
\pagestyle{empty}

```{r label = "num-format", echo =FALSE}
options(scipen=999, digits = 6)
```


\textbf{\LARGE\textsf{Normal distributions and probabilities}}
\vspace{10mm}

<!-- % CA1 (a) -->
\marginnote{\raggedleft \textbf{CA1 (a)}} \vspace{-5mm}

The proportion of Scottish soldiers who had chests measuring between 37 and 42 inches inclusive is 
`r pnorm(42, mean = 40, sd = 2, lower.tail = TRUE) - pnorm(37, mean = 40, sd = 2, lower.tail = TRUE)`.

\vspace{10mm}

<!-- % CA1 (b) -->
\marginnote{\raggedleft \textbf{(b)}} \vspace{-5mm}
The proportion of Scottish soldiers whose chest measurements were greater than 43 inches is 
`r pnorm(43, mean = 40, sd = 2, lower.tail = FALSE)`.

\vspace{10mm}

<!-- % CA1 (c) -->
\marginnote{\raggedleft \textbf{(c)}} \vspace{-5mm}
Only 2.5% of Scottish soldiers had a chest measurement below `r sprintf("%.1f", qnorm(0.025, mean = 40, sd = 2, lower.tail = TRUE))` inches.

\vspace{10mm}

<!-- % CA1 (d) -->
\marginnote{\raggedleft \textbf{(d)}} \vspace{-5mm}
Only 5% of Scottish soldiers had a chest measurement above `r sprintf("%.1f", qnorm(0.95, mean = 40, sd = 2, lower.tail = TRUE))` inches.

\vspace{15mm}

<!-- % CA5 (a) -->
\marginnote{\raggedleft \textbf{CA5 (a)}} \vspace{-5mm}

The proportion of Scottish soldiers had chest measurements more than three standard deviations from the mean, that is, below 34 or above 46 is:
\begin{align*}
  P(X \le 34 + P(X > 46)
    &=  F(34) + (1 - F(46))  \\
    &=  `r pnorm(34, mean = 40, sd = 2, lower.tail = TRUE)`
                              \text{ + (1 - }`r pnorm(46, mean = 40, sd = 2, lower.tail = TRUE)` \text{)} \\
    &= `r pnorm(34, mean = 40, sd = 2, lower.tail = TRUE)`
                              \text{ + } `r (pnorm(46, mean = 40, sd = 2, lower.tail = FALSE))` \\
    &= `r pnorm(34, mean = 40, sd = 2, lower.tail = TRUE) + (1 - pnorm(46, mean = 40, sd = 2, lower.tail = TRUE))`
\end{align*}

\vspace{10mm}

<!-- % CA5 (b) -->
\marginnote{\raggedleft \textbf{(b)}} \vspace{-5mm}
For a normal distribution with mean 0 and variance 2.75 was used to model errors in angular measurements.
According to the model, the proportion of errors are more than three standard deviations from the mean is 
`r 2 * pnorm(0 - 3 * sqrt(2.75), mean = 0, sd = sqrt(2.75))`

OK, we get the message -- \(\mu\) and \(\sigma\) are irrelevant.