---
output:
  pdf_document:
    latex_engine: pdflatex
    keep_tex: true
    template: "o:/dtma/dt-tma.tex"    # replace with "drive:/path/to/template.tex" (my template.tex is available in top folder of M248R repository)

fontsize: 11pt

linestretch: 1.0
---

```{r all_chunks, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r label = "load-pkgs", echo =FALSE, error = FALSE, warnings=FALSE, messages = FALSE}

packages <- c("tidyverse", "kableExtra")
# "tidyverse", "ggrepel", "ggridges", "ggtext", "ggthemes", "ggstatsplot", "ggforce", "rmarkdown", "xaringan", "xaringanthemer", "kableExtra", "shiny", "corrr", "ggdark", "ggsci", "grid", "gridExtra", "showtext", "rcartocolor", "patchwork", "cowplot", "rvest", "jsonlite", "colorspace", "data.table", "latex2exp", "Hmisc", "broom", "reticulate"
# )

xfun::pkg_attach(packages, install = TRUE, message = FALSE)

```

```{r label = "data", echo =FALSE, error = FALSE, warnings=FALSE, messages = FALSE}
options(scipen = 999, digits = 4)

df <- read.csv('o:/M248/data/csvfiles/coins.csv') %>%    # adjust path for source of your data (can be a URL)
  rename(xi = 1) %>% 
  select(xi) %>%
  arrange(xi) %>% 
  mutate(
    i = 1:length(xi), 
    alpha = i / (1 + length(xi)), 
    yi = qnorm(alpha, mean = 0, sd = 1)) %>% 
  relocate(i, .before = xi)

# Extracting the coefficients from the regression equation for the linear model
lmcalc<-lm(yi ~ xi, data = df)

m <- format(signif(unname(coef(lmcalc)[2]), digits = 3), nsmall = 2)
c <- format(signif(unname(coef(lmcalc)[1]), digits = 4), nsmall = 2)
eq_text <- paste("y = ",m,"x + ",c, sep="")
```

\pagestyle{empty}
# QQ plots

\vspace{10mm}

We can compare a sample with a known distribution by using qq plots, which do not require parameters and are easy to interpret.

If we have a random sample of size \(n\) to compare to a standard normal distribution, we can order the data points from smallest to largest to produce \(x_1, x_2, \ldots, x_n\). Then we compare the ordered data points to the expected value of the ordered data points \(q_{1/(n+1)}, q_{2/(n+1)}, \ldots, q_{n/(n+1)}\). Finally, we plot the ordered pairs \(x_n, q_{i/(n+1)}\). This plot is called a \textbf{standard normal plot} or a \textbf{qq plot}.

\vspace{10mm}

\textbf{\large\textsf{Reproducing figure 29 in M248, Book B, page 39:}}
\vspace{5mm}

```{r label = "table", echo = FALSE, error = FALSE, warnings=FALSE, messages = FALSE}
kbl(df, caption = "Table of data and calculations", format = "latex", booktabs = T) %>%
  kable_styling(
    latex_options = c("striped", "hold_position")
    ) %>%
  column_spec(1, width = "1cm") %>%
  column_spec(c(2:3), width = "2cm") %>%
  column_spec(4, width = "2.5cm")
```

\vspace{8mm}

```{r label = 'plot', echo = FALSE, out.width = "80%", fig.asp = 0.6, fig.align = "left", fig.cap = "Normal probability plot"}
ggplot(data = df, 
       aes(
         x = xi, 
         y = yi
       )
  ) +
  geom_point(
    colour = "darkred",
    fill = "white",
    na.rm = TRUE,
    shape = 1,
    size = 2
    ) + 
  geom_smooth(
    method = "lm", 
    formula= y~x, 
    se = FALSE,
    colour = "darkred",
    na.rm = TRUE,
    size = 0.5
    ) +
  annotate(
    "text",
    x = 7.25, 
    y = 1.4, 
    label = eq_text, 
    parse = FALSE
    ) +
  scale_y_continuous(
    limits = c(-2.2, 2.2), 
	   minor_breaks = seq(from = -4, to = 4, by = 0.1)
    ) + 
   scale_x_continuous(
     limits = c(5.5, 8)
     ) +
  theme_bw() +
  labs(
    title="", 
    y="Normal scores",
    x="Silver content of Byzantine coins"
    ) + 
  theme(
    plot.background = element_blank(),  
    panel.background = element_blank(), 
    panel.grid.major.y = element_line(colour = "grey90"),       # element_blank(),
    panel.grid.minor.y = element_line(colour = "grey95"),       # element_blank(), 
    panel.border = element_blank(), 
    axis.line = element_line()
  )
```


